\documentclass[12pt]{article}
%\usepackage[utf8]{inputenc}
\usepackage{amssymb,amsmath,graphicx,url,fullpage,amsfonts,verbatim}
\usepackage[small,bf]{caption}

% add bib to ToC
\usepackage{tocbibind}

\newcommand{\tatonnement}{t\^atonnement}

\input defs.tex

\title{Applications}
\author{A.J. Friend}
\date{\today}

\begin{document}

\maketitle

\section{list of applications}
a list of notes of applications ot hte market stuff

from \cite[p.~104]{nisan2007algorithmic} says that \cite{kelly1997charging}
is a fisher program for TCP congestion problem. for more on these flow problems,
see chapter 22 of the nisan book and maybe chapter 18 on routing games. maybe
section 5.14? and this jain and vazirani paper?
\cite[\S5.13]{nisan2007algorithmic} gives a nice little Fisher application
to resource allocation markets, including a flow market.
\cite[\S5.14]{nisan2007algorithmic} gives a nice example from the previous
section of problem with a single
source and multiple sinks. However, it ends up with edges having zero prices,
so I don't know how that would work in the Arrow-debreu setting, but because
this is a Fisher problem, it should be ok.

\cite[\S7.5]{nisan2007algorithmic} suggests graphical exchange markets, where
agents can only trade with other agents if they share an edge in some graph.
for this to work, we need to allow for local good prices. but everything should
still carry through. cool!

see \cite{gomes2012pure} for a cloud computing market application.

see \cite{vazirani2010spending} for an internet advertising application.

\cite{vazirani2012notion} for throughput allocation problem on a wireless channel


see \cite{srikant2004mathematics,
kelly1997charging,
kelly1998rate,
low1999optimization,
yaiche2000game} for applications to communication network flow control.

\cite{ye2014competitive} communication spectrum market.

\section{places to publish}
\cite{codenotti2008experimental} was published in the acm journal of experimental algorithmics.
where was amin's paper published?
where was the rest of mccune and codenottis stuff published?

ESTEBAN-BRAVO,M. 2004. Computing equilibria in general equilibrium models via interior-point methods. Computational Economic 23, 147–171.
FERRIS,M.C. ANDMUNSON, T. S. 2000a. Complementarity problems in gams and the path solver. Journal

\cite{gomes2012pure} was published in Concurrency and Computation: Practice and Experience

\cite{vazirani2010spending} operations research journal.
\cite{vazirani2012notion} in journal of the ACM.

\bibliographystyle{alpha}
\bibliography{bibliography}

\end{document}